\documentclass{article}

\usepackage{fancyhdr}
\usepackage{extramarks}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{tikz}
\usepackage[plain]{algorithm}
\usepackage{algpseudocode}
\usepackage{enumerate}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{forest}
\usepackage[shortlabels]{enumitem}
\setlist[enumerate, 1]{1\textsuperscript{o}}
\lstset { %
    language=C++,
        backgroundcolor=\color{black!5}, % set backgroundcolor
            basicstyle=\footnotesize,% basic font setting
}

%\usetikzlibrary{automata,positioning}
\usetikzlibrary{positioning,shapes,shadows,arrows,automata}

%
% Basic Document Settings
%

\topmargin=-0.45in
\evensidemargin=0in
\oddsidemargin=0in
\textwidth=6.5in
\textheight=9.0in
\headsep=0.25in

\linespread{1.1}

\pagestyle{fancy}
\lhead{\hmwkAuthorName}
\rhead{ (\hmwkClassInstructor\ \hmwkClassTime): \hmwkTitle}
\lfoot{\lastxmark}
\cfoot{\thepage}

\renewcommand\headrulewidth{0.4pt}
\renewcommand\footrulewidth{0.4pt}

\setlength\parindent{0pt}

\setcounter{secnumdepth}{0}
\newcounter{partCounter}


\newcommand{\hmwkTitle}{Portfolio submission, Topics 7 and 8}
\newcommand{\hmwkDueDate}{May 20, 2016}
\newcommand{\hmwkClass}{CS510 Languages and Low Level Programming}
\newcommand{\hmwkClassTime}{Spring 2016}
\newcommand{\hmwkClassInstructor}{Mark P. Jones}
\newcommand{\hmwkAuthorName}{Konstantin Macarenco}


\title{
    \vspace{2in}
    \textmd{\textbf{\hmwkClass:\ \hmwkTitle}}\\
        \normalsize\vspace{0.1in}\small{Due\ on\ \hmwkDueDate\ at 11:59pm}\\
        \vspace{0.1in}\large{\textit{\hmwkClassInstructor\ \hmwkClassTime}}
    \vspace{3in}
}

\author{\textbf{\hmwkAuthorName}}
\date{}

\renewcommand{\part}[1]{\textbf{\large Part \Alph{partCounter}}\stepcounter{partCounter}\\}

\begin{document}

\maketitle

\pagebreak

%\begin{enumerate}[(a), leftmargin = 0.7cm, nosep]
        \section{Topic 7. Explain the use and implementation of capabilities in access control and resource
        management.}\\

        Capabilities is a method of access control that gives high level of granularity and
        implements ``principle of least privilege''.\\
        Main key points of capabilities:
        \begin{enumerate}[$\bullet$]
            \item Capabilities don't care who owns them.
            \item Capabilities can be delegated (with the same or smaller level of access rights).
            \item Capabilities can be Copied (with the same or smaller level of access rights).
            \item Define specific set of restrictions (same file can be unlocked by two different
                capabilities with different level of access).
        \end{enumerate}

        File descriptors comes very close to this definition, they can be copied, moved to any
        process or completely removed by the system, and defines level of access, i.e. read
        write, or execute.\\
        Although it sounds good in theory this approach is hard to implement in existing
        system - it is incompatible with current implementation of the popular OSes.\\


        Implementation details (taken from seL4):\\
        All capabilities are stored in protected area (within kernel), and not visible from the outside. Each
        process has a capability space (CSpace), and  as system starts up set of capabilities installed to each process,
        according to some policies. Each capability has assigned metadata to store some vital information,
        for example Window Capability might store set of permissions, Untyped Capability has to store pointer
        to the next available address, etc.

        Capabilities management is done by using capability derivation tree, which for performance purposes
        implemented as doubly link list (efficient sequential traverse and quick insert/delete operations).
        This way capabilities can be easily shared with other processes without system need to track each one
        of them.

        When a process requests some action, kernel checks if the process has appropriate capability
        installed in it's CSpace, if not found action is rejected. In other words they can't be
        forged or guessed like Global Thread IDs in L4.

        \section{Topic 8. Develop programs using the capability abstraction provided by the seL4
        microkernel, or capabilities lab.}

        To meet this objective I Completed capability lab, except for the task 8. 
        During the lab I learned basic principles of working with
        capabilities, and confident in understanding them, although real seL4 experience probably would be more valuable.

        In part 7 of the lab ( Improving existing operation) I picked the second approach, even though a bit more
        troublesome to implement, but should be faster. $printString()$ system call, provides address to the region of memory to be
        printed, however this is not safe and should be checked for:
        \begin{enumerate}[1.]
            \item The desired address is not in any way interfere kernel address space. For
            example it can be  a pointer to some area in the 
                kernel space, or a pointer to a not null terminated region of memory that overlaps with kernel's address space.
           \item The Array to be printed must lay within existing memory, it should not point or overlap with any untyped
               regions.
        \end{enumerate}

        During system call kernel finds memory page, where the string belongs and last page
        that can be reached by from that point, check for null in this region.
\end{document}
